---
title: "In-Class Exercise 4"
author: "Yang Tianyi"
editor: visual
execute: 
  echo: true
  eval: true
  warning: false
---

```{r}
pacman::p_load(rstatix, gt, patchwork, tidyverse, webshot2, ggstatsplot)
```

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

![](data:image/png;base64,iVBORw0KGgoAAAANSUhEUgAAACgAAAAaCAYAAADFTB7LAAAAcElEQVR4Xu3OwQmAQAxE0bClWYCW5N06tM6V2YPg5CjoF/JhLoHAi6iqn9eOefUbqrYvHY0cQDLyAlKRNyARmYA0ZMLRkAlGQyaU72tkAtlim7r/vJqDUDjlKBROOQyFU2icQuMUGqfQuBEaV1XPOwEx96nYACK8+wAAAABJRU5ErkJggg== "Run Current Chunk")

```{r}
ggplot(exam_data,
       aes(sample=ENGLISH))+
  stat_qq() +
  stat_qq_line()
```

::: callout-note
We can see that the points deviate significantly from the straight diagonal line. This is a clear indication that the set of data is not normally distributed.
:::

```{r}
qq <- ggplot(exam_data,
       aes(sample=ENGLISH))+
  stat_qq() +
  stat_qq_line()
```

```{r}
sw_t <- exam_data %>%
  shapiro_test(ENGLISH) %>%
  gt()
```

```{r}
tmp <- tempfile(fileext = '.png')
gtsave(sw_t, tmp)
table_png <- png::readPNG(tmp, native = TRUE)
qq + table_png
```

```{r}
set.seed(1234)

gghistostats(
  data = exam_data,
  x = ENGLISH,
  type = "bayes",
  test.value = 60,
  vlab = "English scores"
)
```

```{r}
gghistostats(
  data = exam_data,
  x = ENGLISH,
  type = "p",
  test.value = 60,
  vlab = "English scores"
)
```
