---
title: "Unveiling Demographic and Financial Characteristics of City of Engagement: A Data-driven Study"
subtitle: "Take-home Ex1"
editor: visual
author: Yang Tianyi
execute: 
  warning: false
  message: false
format: html
---

# Setting the Scene

City of Engagement, with a total population of 50,000, is a small city located at Country of Nowhere. The city serves as a service centre of an agriculture region surrounding the city. The main agriculture of the region is fruit farms and vineyards. The local council of the city is in the process of preparing the Local Plan 2023. A sample survey of 1000 representative residents had been conducted to collect data related to their household demographic and spending patterns, among other things. The city aims to use the data to assist with their major community revitalization efforts, including how to allocate a very large city renewal grant they have recently received.

# Getting Started

In the code chunk below, `p_load()` of *pacman* package is used to install and load the following R packages into R environment:

-   tidyverse,

-   lubridate,

-   stringr,

-   ggplot2,

-   knitr,

-   dplyr,

-   gridExtra,

-   plotly,

-   gganimate,

-   cluster

```{r}
pacman::p_load(tidyverse, lubridate, stringr, ggplot2, knitr, dplyr, gridExtra, plotly, gganimate,transformr, cluster)
```

# **The Data**

For the purpose of this study, two data sets are provided. They are:

*Participants.csv*

Contains information about the residents of City of Engagement that have agreed to participate in this study.

-   participantId (integer): unique ID assigned to each participant.

-   householdSize (integer): the number of people in the participant's household

-   haveKids (boolean): whether there are children living in the participant's household.

-   age (integer): participant's age in years at the start of the study.

-   educationLevel (string factor): the participant's education level, one of: {"Low", "HighSchoolOrCollege", "Bachelors", "Graduate"}

-   interestGroup (char): a char representing the participant's stated primary interest group, one of {"A", "B", "C", "D", "E", "F", "G", "H", "I", "J"}. Note: specific topics of interest have been redacted to avoid bias.

-   joviality (float): a value ranging from \[0,1\] indicating the participant's overall happiness level at the start of the study.

*FinancialJournal.csv*

Contains information about financial transactions.

-   participantId (integer): unique ID corresponding to the participant affected

-   timestamp (datetime): the time when the check-in was logged

-   category (string factor): a string describing the expense category, one of {"Education", "Food", "Recreation", "RentAdjustment", "Shelter", "Wage"}

-   amount (double): the amount of the transaction

```{r}
Journal <- read_csv("data/FinancialJournal.csv")
```

```{r}
participants <- read_csv("data/Participants.csv")
```

## Data Cleaning

First, we will have to start with missing value check on both datasets.

```{r}
# Missing value check for financial_journal dataset
missing_values_fj <- sapply(Journal, function(x) sum(is.na(x)))
print(missing_values_fj)

# Missing value check for participants dataset
missing_values_participants <- sapply(participants, function(x) sum(is.na(x)))
print(missing_values_participants)

```

Seems no missing value appears in the dataset, we them move on to the next part.

Next part is checking for duplicate participant IDs: Ensure that each participant ID is unique and there are no duplicates. I use the **`duplicated()`** function to identify and remove any duplicate rows.

```{r}
# Checking for duplicate participant IDs
duplicate_ids <- participants[duplicated(participants$participantId), "participantId"]
if (length(duplicate_ids) > 0) {
  print(paste("Duplicate participant IDs found:", paste(duplicate_ids, collapse = ", ")))
} else {
  print("No duplicate participant IDs found.")
}
```

Next, I will perform categorical variable standardization, the levels of categorical variables like "educationLevel" or "interestGroup" might inconsistent capitalization or spelling, I decided to use functions like **`tolower()`** or **`recode()`** from the **`dplyr`** to standarde them for consistency and easier analysis.

```{r}
library(dplyr)

# Standardizing the educationLevel variable
participants <- participants %>%
  mutate(educationLevel = recode(educationLevel,
                                 "Low" = "Low",
                                 "HighSchoolOrCollege" = "High School or College",
                                 "Bachelors" = "Bachelors",
                                 "Graduate" = "Graduate"))

# Standardizing the interestGroup variable
participants <- participants %>%
  mutate(interestGroup = recode(interestGroup,
                                "A" = "Group A",
                                "B" = "Group B",
                                "C" = "Group C",
                                "D" = "Group D",
                                "E" = "Group E",
                                "F" = "Group F",
                                "G" = "Group G",
                                "H" = "Group H",
                                "I" = "Group I",
                                "J" = "Group J"))

participants <- participants %>%
  mutate(haveKids = if_else(haveKids == "TRUE", TRUE, FALSE)) # Convert "TRUE" and "FALSE" to boolean TRUE and FALSE
```

Next, we need to validate the "timestamp" variable in the "financial_journal" dataset, so I converted the "timestamp" column to a proper datetime format

```{r}
Journal$timestamp <- as.POSIXct(Journal$timestamp, format = "%Y-%m-%d %H:%M:%S")
```

Now all the seems to be cleaned, lets take a quick look at each table:

```{r}
head(Journal)
```

```{r}
head(participants)
```

Now that the data cleaning steps have been completed, we can proceed with Exploratory Data Analysis (EDA)

# Exploratory Data Analysis

EDA involves analyzing and visualizing the data to gain insights, understand patterns, and uncover relationships among variables.

## Descriptive statistics

To compute descriptive statistics for numerical variables in R, we can use the **`summary()`** function.

Here's how I create plots for the variables in the **`participants`** dataframe and the **`Journal`** dataframe using the **`ggplot2`** package in R:

```{r}
# Histogram for age
histogram_age <- ggplot(participants, aes(x = age)) +
  geom_histogram(fill = "lightblue", color = "black") +
  labs(title = "Distribution of Age", x = "Age", y = "Count")

# Bar plot for educationLevel
barplot_education <- ggplot(participants, aes(x = educationLevel)) +
  geom_bar(fill = "lightblue", color = "black") +
  labs(title = "Distribution of Education Level", x = "Education Level", y = "Count")

# Bar plot for interestGroup
barplot_interest <- ggplot(participants, aes(x = interestGroup)) +
  geom_bar(fill = "lightblue", color = "black") +
  labs(title = "Distribution of Interest Group", x = "Interest Group", y = "Count")

# Calculate median joviality for each age group
median_joviality <- participants %>%
  group_by(age) %>%
  summarize(median_joviality = median(joviality))

# Plot median joviality by age group
scatterplot_age_joviality <- ggplot(median_joviality, aes(x = age, y = median_joviality)) +
  geom_line() +
  labs(title = "Median Joviality by Age Group", x = "Age", y = "Median Joviality")

# Arrange plots in a grid
grid.arrange(histogram_age, barplot_education, barplot_interest, scatterplot_age_joviality,
             nrow = 2, ncol = 2)
```

For the Age, the distribution seems very rather smooth with no clear peak or hump of a group age, it jumps up and down in between ages with no concentrated point.

For the Education level, people who have high school or college level appears to be most common.

For the interest group and median joviality, either of them appear to have any normal distribution symptom.

```{r}
# Box plot for amount
boxplot_amount <- ggplot(Journal, aes(x = "", y = amount)) +
  geom_boxplot(fill = "lightblue", color = "black", outlier.shape = NA) +
  coord_cartesian(ylim = c(-10, 50)) +  # Set the y-axis limits as desired
  labs(title = "Distribution of Amount", x = "", y = "Amount")

# Bar plot for category
barplot_category <- ggplot(Journal, aes(x = category)) +
  geom_bar(fill = "lightblue", color = "black") +
  labs(title = "Distribution of Expense Categories", x = "Category", y = "Count")

# Plot total amount spent per month
Journal$month <- floor_date(Journal$timestamp, "month")
monthly_totals <- Journal %>%
  group_by(month) %>%
  summarize(total_amount = sum(amount))

lineplot_total_amount <- ggplot(monthly_totals, aes(x = month, y = total_amount)) +
  geom_line() +
  labs(title = "Total Amount Spent per Month", x = "Month", y = "Total Amount")

# Arrange plots in a grid
grid.arrange(boxplot_amount, barplot_category, lineplot_total_amount,
             nrow = 2, ncol = 2)
```

For the amount, as it concentrated towards the 0, it means the amount of transactions are mostly small.

For the Expense category, the transaction categories are mostly on food, recreation, and wages.

For the total amount spent per month, except for before April 2022, all other time spam seems quite smooth with transaction amount.

## Relationships between variables

After finishing the descriptive statistics, let's move on to exploring the relationships between variables. Understanding the relationships between variables can provide valuable insights into the data and help uncover interesting patterns or associations.

In order for a better comparison and visualization in between variables, we should join the dtaframe:

```{r}
# Join the participants and financial_journal dataframes
merged_data <- participants %>%
  inner_join(Journal, by = "participantId")

# Check for missing values in the age variable
missing_age <- sum(is.na(merged_data$age))

# Print the number of missing values
print(paste("Number of missing values in age:", missing_age))

# Filter out the rows with missing values in age
merged_data <- merged_data[!is.na(merged_data$age), ]
```

### Age and Joviality

```{r}
# Define the age groups
age_groups <- c("18-25", "26-35", "36-45", "46-55", "56+")

# Create a new variable for age groups
merged_data <- merged_data %>%
  mutate(age_group = cut(age, breaks = c(18, 25, 35, 45, 55, Inf), labels = age_groups))

# Box plot for age groups and joviality
ggplot(merged_data, aes(x = age_group, y = joviality)) +
  geom_boxplot() +
  labs(title = "Joviality by Age Group", x = "Age Group", y = "Joviality")

# Calculate summary statistics by age group
summary_age <- merged_data %>%
  group_by(age_group) %>%
  summarize(
    mean_joviality = mean(joviality),
    median_joviality = median(joviality),
    min_joviality = min(joviality),
    max_joviality = max(joviality)
  )

# Print the summary statistics
print(summary_age)
```

Based on the summary statistics for the age groups and joviality, we can interpret the results as follows:

-   Age Group "18-25":

    -   Mean Joviality: 0.511

    -   Median Joviality: 0.503

    -   Minimum Joviality: 0.001

    -   Maximum Joviality: 0.986

-   Age Group "26-35":

    -   Mean Joviality: 0.508

    -   Median Joviality: 0.504

    -   Minimum Joviality: 0.001

    -   Maximum Joviality: 0.999

-   Age Group "36-45":

    -   Mean Joviality: 0.516

    -   Median Joviality: 0.521

    -   Minimum Joviality: 0.002

    -   Maximum Joviality: 0.998

-   Age Group "46-55":

    -   Mean Joviality: 0.440

    -   Median Joviality: 0.429

    -   Minimum Joviality: 0.000

    -   Maximum Joviality: 0.999

-   Age Group "56+":

    -   Mean Joviality: 0.510

    -   Median Joviality: 0.491

    -   Minimum Joviality: 0.000

    -   Maximum Joviality: 0.992

These statistics provide an overview of the joviality levels within each age group. Here are some key observations:

-   The mean joviality levels are relatively consistent across the age groups, ranging from approximately 0.440 to 0.516.

-   The median joviality levels show a similar pattern, with values ranging from around 0.429 to 0.521.

-   The minimum and maximum joviality levels vary within each age group, indicating the range of joviality experienced by individuals in each group.

-   Overall, there doesn't seem to be a clear trend or pattern in joviality levels across age groups based on these summary statistics.

### Household Size and Education Level

```{r}
# Bar plot for household size and education level
ggplot(merged_data, aes(x = householdSize, fill = educationLevel)) +
  geom_bar() +
  labs(title = "Household Size by Education Level", x = "Household Size", y = "Count") +
  scale_fill_discrete(name = "Education Level")
```

From the plot above, it appears that the majority of households in the dataset have a household size of 1 or 2. Additionally, the "High School or College" education level seems to be the most prevalent across all household sizes.

This suggests a possible correlation between household size and education level. It is likely that individuals living alone or in smaller households (size 1 or 2) are more likely to have completed high school or attended college, as reflected by the higher count in the "High School or College" category for these household sizes.

On the other hand, larger household sizes (beyond size 2) may include a mix of different education levels, as evidenced by the presence of "Bachelors" and "Graduate" categories in these groups. This suggests a less diversity of education levels within larger households.

### Education Level and Joviality

```{r}
# Box plot for education level and joviality
ggplot(merged_data, aes(x = educationLevel, y = joviality)) +
  geom_boxplot() +
  labs(title = "Joviality by Education Level", x = "Education Level", y = "Joviality")
```

Based on these results, we can make a scientific guess that individuals with higher education levels, such as "Graduate" and "Bachelors," tend to have higher median joviality scores compared to those with lower education levels, such as "Low" and "High School or College." Additionally, individuals with higher education levels generally have a wider range of joviality values, as indicated by the larger differences between the minimum and maximum joviality values.

### Interest Group and Joviality

```{r}
# Box plot for interest group and joviality
ggplot(merged_data, aes(x = interestGroup, y = joviality)) +
  geom_boxplot() +
  labs(title = "Joviality by Interest Group", x = "Interest Group", y = "Joviality")
```

Based on these results, we can make a scientific guess that different interest groups may have varying levels of median joviality. For example, "Group E" and "Group B" have relatively higher median joviality compared to "Group H" and "Group D." However, it's important to note that there could be other factors influencing joviality, and this analysis only considers the provided data.

### Household Size and Amount

```{r}
# Scatter plot for household size and amount
ggplot(merged_data, aes(x = as.factor(householdSize), y = amount, fill = as.factor(householdSize))) +
  geom_bar(stat = "summary", fun = "mean", position = "dodge") +
  labs(title = "Mean Amount by Household Size", x = "Household Size", y = "Mean Amount") +
  scale_fill_discrete(name = "Household Size")
```

Based on these results, we can make a scientific analysis that households with larger sizes tend to have slightly higher mean amounts compared to smaller households. However, it's important to note that the median amount is the same for all household sizes, suggesting a similar distribution of amounts regardless of household size.

Additionally, the range of amounts spent varies across different household sizes, as indicated by the differences in the minimum and maximum amounts. This implies that there might be variations in spending habits and financial situations among households of different sizes.

### Education Level and Amount

```{r}
# Box plot for education level and amount
ggplot(merged_data, aes(x = educationLevel, y = amount, fill = educationLevel)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(color = educationLevel), width = 0.2, height = 0, alpha = 0.5) +
  labs(title = "Distribution of Amount by Education Level", x = "Education Level", y = "Amount") +
  scale_fill_discrete(name = "Education Level") +
  scale_color_discrete(name = "Education Level")
```

Based on these results, we can make a scientific analysis that individuals with higher education levels, such as "Graduate" and "Bachelors," tend to have higher maximum amounts spent compared to those with lower education levels, such as "Low" and "High School or College."

The data reveals that within each education level group, there are individuals who have higher amounts spent or higher income compared to others in the same group. This suggests that being successful is always a possibility, regardless of education level, as personal circumstances, career choices, skills, networking, and opportunities also play a significant role.Education level is an important factor that can contribute to success and income, but it is not the sole determinant. Education provides knowledge and opportunities, but it does not guarantee automatic financial prosperity. Success is multifaceted and influenced by various factors beyond education alone.

### Interest Group and Amount

```{r}
# Box plot for interest group and amount
ggplot(merged_data, aes(x = interestGroup, y = amount, fill = interestGroup)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(color = interestGroup), width = 0.2, height = 0, alpha = 0.5) +
  labs(title = "Distribution of Amount by Interest Group", x = "Interest Group", y = "Amount") +
  scale_fill_discrete(name = "Interest Group") +
  scale_color_discrete(name = "Interest Group")
```

There appears no distinct and big differences across different interest groups, except fir B, C, D, E, G and J, there is a small chunk of them that plots right above the majority, indicating that there might be a specific class within the interest groups.

### Age and Amount

```{r}
# Scatter plot for age and amount
ggplot(merged_data, aes(x = age_group, y = amount, fill = age_group)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(color = age_group), width = 0.2, height = 0, alpha = 0.5) +
  labs(title = "Distribution of Amount by Age Group", x = "Age Group", y = "Amount") +
  scale_fill_discrete(name = "Age Group") +
  scale_color_discrete(name = "Age Group")
```

For the age group and the expenses plot, it appears to have a similar pattern with the previous two plots. While distinct classes are found, indicates age might not be de determinant factor of different class.

Analyzing the relationship between age group and financial outcomes is complex and multidimensional. It requires considering a broader range of variables and conducting more in-depth statistical analyses to gain a comprehensive understanding of the factors influencing success and income.

### Education Level and Expenses Category

```{r}
# Bar plot for education level and expense category (mean expenses)
ggplot(merged_data, aes(x = educationLevel, y = amount, fill = category)) +
  stat_summary(fun = "mean", geom = "bar", position = "dodge") +
  labs(title = "Mean Expense Category by Education Level", x = "Education Level", y = "Mean Expense") +
  scale_fill_discrete(name = "Expense Category")
```

For individuals with a Graduate level of education, the mean amount spent on wages appears to be the highest among all expense categories. Additionally, the highest mean spending in the "Shelter" category suggests that individuals with a Graduate degree allocate a significant portion of their expenses towards their living arrangements.

For individuals with a Bachelor's degree, the "RentAdjustment" category tends to have the highest mean amount compared to other expense categories. This indicates that individuals with a high level of education, such as a Bachelor's degree, allocate a larger proportion of their expenses towards adjusting their rent.

On the other hand, for individuals with a low education level, the mean spending on shelter is the lowest among all groups. This implies that individuals with lower education levels tend to allocate less money towards their living environment or housing expenses. In other words, lower education levels may play a role in determining lower spending on shelter, indicating potential differences in financial capabilities and living conditions based on educational attainment.

### Interest Group and Expense Category

```{r}
# Bar plot for interest group and expense category (mean values)
ggplot(merged_data, aes(x = interestGroup, fill = category)) +
  stat_summary(aes(y = amount, group = category), fun = "mean", geom = "bar") +
  labs(title = "Mean Expense Category by Interest Group", x = "Interest Group", y = "Mean Count") +
  scale_fill_discrete(name = "Expense Category")
```

For interest groups E, G, and I, the mean count for the "RentAdjustment" category appears to be relatively low compared to other interest groups. This suggests that individuals in these groups, on average, spend less on adjusting their rent compared to individuals in other interest groups.

Conversely, interest group H shows a significant difference in renting behavior. The mean count for the "RentAdjustment" category is notably higher in group H compared to other interest groups. This indicates that individuals in interest group H, on average, allocate a larger portion of their expenses towards adjusting their rent.

For other expense categories, such as "Education," "Food," "Recreation," and others, there are no significant differences observed across the interest groups. The mean counts for these categories appear to be relatively similar, suggesting that individuals in different interest groups have comparable spending patterns in these areas.

# Dynamic Visualizations

This section presents visually engaging and animated plots that dynamically illustrate patterns and trends in the data. These dynamic visualizations enhance the understanding of complex relationships and temporal changes by animating data points or incorporating motion. Through these interactive and dynamic elements, viewers can explore the data in a more engaging and intuitive way, enabling them to grasp insights and patterns that may not be readily apparent in static visualizations. This section adds an element of interactivity and storytelling to the data analysis process, facilitating effective communication and interpretation of the findings.

## Age vs. Joviality

```{r}
# Create an interactive scatter plot with tooltips
plot1 <- plot_ly(data = participants, x = ~age, y = ~joviality, color = ~educationLevel,
             text = ~paste("Participant ID:", participantId, "<br>",
                           "Education Level:", educationLevel, "<br>",
                           "Interest Group:", interestGroup),
             hoverinfo = "text",
             type = "scatter", mode = "markers") %>%
  layout(title = "Interactive Scatter Plot: Age vs. Joviality",
         xaxis = list(title = "Age"),
         yaxis = list(title = "Joviality"))

# Display the interactive plot
plot1
```

The resulting plot allows users to interact with the data points by hovering over them to view additional information. The scatter plot visualizes the relationship between age and joviality, with each data point represented by a marker color-coded based on the education level. This interactive plot provides a dynamic way to explore and analyze the data.

## Joviality over Time

```{r}
plot2 <- ggplot(merged_data, aes(x = timestamp, y = joviality, color = educationLevel, size = householdSize, shape = interestGroup)) +
  geom_point() +
  scale_color_manual(values = c("#FF0000", "#00FF00", "#0000FF", "#FFFF00"), name = "Education Level") +
  scale_size(range = c(2, 10), name = "Household Size") +
  scale_shape_manual(values = c(16, 17, 18, 19, 20, 21, 22, 23, 24, 25), name = "Interest Group") +
  labs(title = "Joviality over Time", x = "Timestamp", y = "Joviality") +
  transition_time(timestamp) +
  ease_aes('linear')

# Animate and save as a GIF
anim1 <- animate(plot2, nframes = 100)
anim_save("animated_plot1.gif", anim1, fps = 20)
anim1
```

The animated plot shows the joviality levels over time for different individuals in the dataset. Each point represents a participant, and their positions on the plot indicate their joviality level at a specific timestamp. The color of the points represents the education level of the participants, the size represents their household size, and the shape represents their interest group.

As the animation progresses, we can observe how the joviality levels change over time for individuals with different education levels, household sizes, and interest groups. The plot allows us to identify any patterns, trends, or fluctuations in joviality over time, and how they might differ across the various demographic groups.

By combining the timestamp, joviality, education level, household size, and interest group in an interactive and animated visualization, we can gain insights into the dynamics of joviality within different segments of the population and explore any potential relationships or correlations between these variables.

## Expenses and Food Over Time

```{r}
# Convert timestamp to date format
merged_data$date <- as.Date(merged_data$timestamp)

# Filter expenses and wages data
expenses <- merged_data[merged_data$category != "Food", ]
food <- merged_data[merged_data$category == "Food", ]

# Create a line plot of total expenses over time
plot3 <- ggplot() +
  geom_line(data = expenses, aes(x = date, y = amount, color = "Expenses"), size = 1) +
  geom_line(data = food, aes(x = date, y = amount, color = "Food"), size = 3) +
  labs(title = "Expenses and Food Over Time", x = "Date", y = "Amount", color = "Category") +
  scale_color_manual(values = c("Expenses" = "blue", "Food" = "green")) +
  coord_cartesian(ylim = c(-100, 1000)) +  # Adjust the y-axis limit
  transition_time(date) +
  ease_aes("linear")

# Animate the plot
anim2 <- animate(plot3, nframes = 200, duration = 10)
anim_save("animated_plot2.gif", anim2, fps = 10)
anim2
```

This plot displays the trends in expenses and food expenses over a specific time period. The x-axis represents the date, while the y-axis represents the amount of expenses. The plot includes two lines: one representing total expenses and the other representing food expenses.

The blue line represents the total expenses, showing the overall trend in expenses over time. The green line represents food expenses, specifically highlighting the trend in spending on food items. By comparing the two lines, we can observe how food expenses contribute to the overall expenses and whether there are any notable patterns or fluctuations.

The visualization shows the distribution of expenses over time, with a specific focus on the food category. It is evident that food expenses constitute a relatively small portion of the overall expenses. This observation suggests that individuals in the town allocate a smaller proportion of their budget towards food compared to other expenditure categories.

# Cluster Analysis

This section focuses on identifying distinct groups or clusters within the dataset based on similarities in the selected variables. This analysis helps uncover hidden patterns or structures in the data by grouping similar observations together and highlighting their shared characteristics. By applying clustering algorithms, such as k-means or hierarchical clustering, the data is partitioned into clusters, with each cluster representing a distinct segment or subgroup within the population. The cluster analysis provides valuable insights into the heterogeneity within the data and can be used for segmentation, profiling, or targeted strategies. By understanding the unique characteristics and behaviors of each cluster, decision-makers can tailor their approaches and interventions to specific groups, leading to more effective decision-making and resource allocation.

## Age, Household Size, and Spending Amount

K-means clustering was chosen as the method for analyzing the variables age, household size, and spending amount. This approach provides a straightforward and efficient way to identify distinct groups based on similarities in these variables. By clustering individuals based on age, we can explore if there are any age-related patterns in household size and spending amount. The clustering of household size allows us to examine if different household sizes exhibit unique spending behaviors. Additionally, clustering individuals based on spending amount helps identify distinct spending profiles or segments within the population. Overall, the application of k-means clustering on age, household size, and spending amount enables us to uncover underlying patterns and relationships, assisting in targeted decision-making and strategy development based on different age groups, household sizes, and spending patterns.

```{r}
set.seed(123)  # Set a specific seed value

# Perform cluster analysis using k-means
data_for_kmeans <- merged_data %>%
  select(age, householdSize, amount)

# Standardize the variables
scaled_data <- scale(data_for_kmeans)

# Determine the optimal number of clusters using the elbow method
wss <- numeric(10)
for (k in 1:10) {
  model <- kmeans(scaled_data, centers = k)
  wss[k] <- sum(model$withinss)
}

# Perform k-means clustering with the chosen number of clusters
k <- 3  # Choose the number of clusters based on the elbow plot
kmeans_model <- kmeans(scaled_data, centers = k)

# Add the cluster assignments to the original dataset
merged_data$Kmeans_cluster <- as.factor(kmeans_model$cluster)

# Calculate cluster means
Kmeans_summary <- merged_data %>%
  group_by(Kmeans_cluster) %>%
  summarize(mean_age = mean(age),
            mean_householdSize = mean(householdSize),
            mean_amount = mean(amount))

# Print the cluster summary
print(Kmeans_summary)
```

\
Cluster 1:

Mean Age: 34.53

Mean Household Size: 2.61

Mean Amount: -2.13

Cluster 1: This cluster has a relatively younger population with a mean age of 34.53. The average household size is 2.61, suggesting that individuals in this cluster tend to live in larger households. The mean amount is -2.13, indicating that on average, individuals in this cluster have lower spending amounts.

This cluster represents relatively younger individuals who live in larger households. They have lower spending amounts compared to the other clusters. This group may include young adults or families who are more cautious with their expenses and prioritize saving.

Cluster 2:

Mean Age: 40.43

Mean Household Size: 2.20

Mean Amount: 212.26

Cluster 2: This cluster has a slightly older population with a mean age of 40.43. The average household size is 2.20, suggesting that individuals in this cluster tend to live in smaller households. The mean amount is 212.26, indicating that on average, individuals in this cluster have higher spending amounts.

This cluster consists of slightly older individuals who live in smaller households. They have higher spending amounts, suggesting that they may have higher incomes or different spending priorities. This group might include established professionals or affluent individuals who can afford to spend more on various expenses.

Cluster 3:

Mean Age: 42.71

Mean Household Size: 1.26

Mean Amount: 6.49

Cluster 3: This cluster has an older population with a mean age of 42.71. The average household size is 1.26, indicating that individuals in this cluster tend to live alone or in smaller households. The mean amount is 6.49, suggesting that on average, individuals in this cluster have moderate spending amounts.

This cluster represents older individuals who live alone or in smaller households. They have moderate spending amounts, indicating a balanced approach to expenses. This group could include retirees or individuals who have a more frugal lifestyle, focusing on essential needs rather than extravagant spending.

## Age, Household Size, Joviality and Education Level

The fanny algorithm was selected for analyzing the variables of age, household size, joviality, and education level. Fanny, a variant of k-means clustering, is particularly suitable for categorical variables such as education level. By incorporating education level as a categorical variable, we can explore the relationship between demographic factors (age and household size) and psychological well-being (joviality) within different educational groups. Fanny algorithm allows for a more flexible clustering approach by accommodating mixed variable types, enabling us to uncover distinct clusters based on these diverse dimensions. This approach provides a comprehensive understanding of how age, household size, joviality, and education level interrelate, allowing for targeted interventions and tailored strategies to address specific needs within different demographic and educational segments of the population.

Given the high computational requirements of cluster analysis, a subsample of 5000 observations was chosen to expedite the analysis process. By working with a smaller subset, the computational time was significantly reduced while still allowing for meaningful insights. It is important to note that the subsample should be representative of the larger dataset, capturing its key characteristics and variability. The use of a subsample should be clearly documented and acknowledged to ensure transparency in the analysis.

```{r}
set.seed(1234)  # Set a specific seed value

# Subsample the dataset
sample_size <- 5000  # Adjust the sample size as needed
subsample <- merged_data %>% sample_n(sample_size)

# Select variables for cluster analysis
data_for_fanny <- subsample %>%
  select(age, householdSize, joviality, educationLevel)

# Convert education level to numeric labels
data_for_fanny$educationLabel <- as.factor(data_for_fanny$educationLevel)

# Standardize the numeric variables
scaled_data <- scale(data_for_fanny[, c("age", "householdSize", "joviality")])

# Perform clustering with Fanny algorithm
k <- 4  # Number of clusters
fanny_model <- fanny(scaled_data, k, memb.exp = 1.1)

# Add the cluster assignments to the subsample dataset
subsample$cluster <- as.factor(fanny_model$clustering)

# Calculate cluster means
fanny_summary <- subsample %>%
  group_by(cluster) %>%
  summarize(mean_age = mean(age),
            mean_householdSize = mean(householdSize),
            mean_joviality = mean(joviality),
            mean_educationLevel = levels(data_for_fanny$educationLabel)[which.max(table(data_for_fanny$educationLabel))])

# Print the cluster summary
print(fanny_summary)
```

Cluster 1:

Mean Age: 30.57

Mean Household Size: 1.51

Mean Joviality: 0.799

Mean Education Level: High School or College

Cluster 1: This cluster represents individuals with a mean age of 30.57, relatively smaller household sizes of 1.51, and a higher level of joviality with a mean of 0.799. The majority of individuals in this cluster have a high school or college education level.

This group might include young adults who are generally cheerful and have smaller households. They could be at a stage in life where they are focused on personal growth, career development, and building relationships. Their higher level of joviality might be influenced by their youthful energy and optimism.

Cluster 2:

Mean Age: 51.67

Mean Household Size: 1.47

Mean Joviality: 0.463

Mean Education Level: High School or College

Cluster 2: This cluster consists of individuals with a mean age of 51.67 and relatively smaller household sizes of 1.47. They have a lower level of joviality with a mean of 0.463. Similar to other clusters, the majority of individuals in this cluster also have a high school or college education level.

This group might represent older individuals who tend to be less jovial but have smaller households. They might be in a stage of life where they have more responsibilities, such as taking care of family or managing their careers. Their lower level of joviality might be influenced by life experiences and responsibilities.

Cluster 3:

Mean Age: 30.06

Mean Household Size: 1.56

Mean Joviality: 0.254

Mean Education Level: High School or College

Cluster 3: This cluster represents individuals with a mean age of 30.06 and slightly larger household sizes of 1.56. They have a lower level of joviality with a mean of 0.254. Similar to the other clusters, the majority of individuals in this cluster have a high school or college education level.

This group might include younger individuals with slightly larger households but lower joviality levels. They could be in a stage of life where they are starting families or building their careers. Their lower level of joviality might be influenced by the challenges and responsibilities they face as they transition into adulthood.

Cluster 4:

Mean Age: 38.31

Mean Household Size: 3.00

Mean Joviality: 0.477

Mean Education Level: High School or College

Cluster 4: This cluster consists of individuals with a mean age of 38.31 and larger household sizes of 3.00. They have a moderate level of joviality with a mean of 0.477. Similar to other clusters, the majority of individuals in this cluster have a high school or college education level.

This group might represent individuals with larger families who tend to be moderately jovial. They could be in a stage of life where they are actively managing their family life, balancing multiple responsibilities, and engaging in social interactions. Their moderate level of joviality might be influenced by their family dynamics and social connections.

## Summary

Based on the cluster analysis findings, we have identified distinct groups within the dataset based on demographic and behavioral characteristics. The analysis revealed variations in age, household size, spending patterns, and education level among the clusters. These findings provide valuable insights into the different segments within the dataset and can be used to understand the diversity among individuals in terms of their financial behaviors and characteristics. The identified clusters can help inform targeted marketing strategies, personalized financial planning, and decision-making processes to better meet the specific needs and preferences of individuals within each cluster. Overall, the cluster analysis has provided a deeper understanding of the dataset and offers valuable information for making data-driven business and financial decisions.

# Conclusion

In conclusion, the analysis of the data has provided valuable insights for the city's community revitalization efforts and the allocation of the city renewal grant. The cluster analysis revealed distinct groups within the population based on age, household size, education level, and spending behavior. These findings shed light on various aspects of the community and can inform strategic resource allocation and decision-making processes.

1.  Age Group: The age group of 46-55 appears to have lower joviality compared to other age groups. This finding suggests that individuals in this age range might exhibit a lower level of happiness or cheerfulness, which could be influenced by various factors such as life circumstances or personal experiences.

2.  Household Size and Education Level: The proportion of individuals with a "low" education level is higher among those living in households with a size of 1. This observation indicates that smaller households are more likely to have individuals with a lower level of education. This relationship could be influenced by factors such as individual preferences, financial constraints, or life circumstances.

3.  Education Level and Joviality: Individuals with a "low" education level exhibit a lower median joviality level. This finding suggests a potential association between education level and emotional well-being, indicating that individuals with lower education levels might experience lower levels of joviality on average. However, it's important to consider that individual differences and external factors may also contribute to variations in joviality levels.

4.  Education Level and Spending on Recreation: Individuals with a bachelor's level of education tend to spend more on recreation activities. This finding suggests that higher education levels might be associated with a greater interest in leisure and recreational pursuits. This could be influenced by factors such as higher disposable income, access to educational opportunities that promote recreational interests, or personal preferences.

5.  Education Level and Maximum Spending Amount: Individuals with a graduate-level education exhibit a higher maximum spending amount. This finding indicates that individuals with higher education levels might have higher earning potentials or financial resources, allowing them to engage in higher levels of spending. This could be influenced by factors such as career advancements, higher-paying job opportunities, or greater financial stability associated with higher education levels.

6.  K-means analysis: We identified three distinct clusters. Cluster 1 consists of individuals with a relatively younger age (mean age: 34.53) and moderate household size (mean household size: 2.61). However, they exhibit a negative mean spending amount (-2.13), indicating potential financial challenges. Cluster 2 represents a slightly older group (mean age: 40.43) with a similar household size (mean household size: 2.20), but they display a significantly higher mean spending amount (212.26). This suggests that they may have higher disposable income or engage in higher-value transactions. Cluster 3 represents a relatively older group (mean age: 42.71) with a smaller household size (mean household size: 1.26) and a modest mean spending amount (6.49).

7.  Fanny analysis: We identified four distinct clusters. Cluster 1 consists of younger individuals (mean age: 30.57) with a moderate household size (mean household size: 1.51) and relatively high joviality (mean joviality: 0.80). The majority of individuals in this cluster have a high school or college education level. Cluster 2 represents an older group (mean age: 51.67) with a similar household size (mean household size: 1.47) but lower joviality (mean joviality: 0.46). Cluster 3 consists of younger individuals (mean age: 30.06) with a slightly larger household size (mean household size: 1.56) and lower joviality (mean joviality: 0.25). Cluster 4 represents individuals with a moderate age (mean age: 38.31), larger household size (mean household size: 3.00), and moderate joviality (mean joviality: 0.48). The majority of individuals in all clusters have a high school or college education level.These findings provide actionable insights for the city's decision-makers, enabling them to develop targeted programs and interventions that address the specific needs and preferences of different segments of the population. By leveraging these insights, the city can make informed decisions regarding the allocation of resources, ensuring that the community's needs are effectively addressed and that the city renewal grant is utilized for maximum impact.

The analysis of the data has provided valuable insights into the community's demographics, spending behavior, and educational levels, revealing distinct patterns and associations. These findings have shed light on various aspects of the community and highlighted areas where targeted interventions and resource allocation can make a meaningful impact. Based on these findings, several actionable recommendations emerge.

1.  Targeted Support for Age Group with Lower Joviality: Develop programs and initiatives that promote emotional well-being and enhance the joviality of the age group (46-55) exhibiting lower levels of happiness. This could include mental health support, social engagement activities, or community events aimed at fostering a sense of joy and contentment.

2.  Educational Opportunities for Smaller Households and Lower Education Levels: Allocate resources to provide educational opportunities, vocational training, or skill development programs for individuals living in smaller households and those with lower education levels. This can empower them with knowledge and skills, enhancing their employability and long-term prospects.

3.  Investment in Recreational Facilities and Activities: Direct resources towards the development of recreational facilities, community centers, parks, and cultural programs that cater to the interests and preferences of individuals with bachelor's degrees. This can provide them with enriching leisure experiences and promote a vibrant community life.

4.  Support Entrepreneurship and Economic Empowerment for Graduate-Level Individuals: Allocate resources to support entrepreneurship programs, mentorship initiatives, and business development opportunities for individuals with graduate-level education. This can stimulate economic growth, job creation, and innovation within the community.

5.  Tailored Financial and Business Support for Different Clusters: Provide targeted financial and business support to different clusters based on their characteristics. For example, Cluster 1, consisting of younger individuals with moderate household size, could benefit from financial literacy programs and entrepreneurship training. Cluster 2, with older individuals and higher spending amounts, might benefit from financial planning and retirement savings initiatives. Cluster 3, with a larger household size, could benefit from support programs for family finances and budgeting. Cluster 4, with a moderate household size and joviality, could benefit from programs that enhance financial management skills and promote savings.

6.  Collaboration and Partnerships: Foster collaborations and partnerships with local educational institutions, community organizations, and businesses to leverage their expertise, resources, and networks. This can enhance the effectiveness of resource allocation efforts and create synergies for community revitalization.

At the end, the comprehensive analysis of the data, including exploratory data analysis, interactive visualizations, and cluster analysis, has provided valuable insights for the city's community revitalization efforts and the allocation of the city renewal grant. By uncovering distinct demographic groups, spending behavior patterns, and educational associations, we have gained a deeper understanding of the community's needs, preferences, and potential areas of improvement. Through the implementation of targeted actions, such as supporting specific age groups, promoting educational opportunities, investing in recreational facilities, and fostering entrepreneurship, the city can harness these insights to drive positive change and create a vibrant and inclusive community. The findings from this analysis, coupled with the proposed actions, form a solid foundation for strategic decision-making and resource allocation, positioning the city for a successful and impactful community revitalization journey.
